Fixing Vulnerabilities Hinders Maintainability

### 2019-01-30 paper/ICPC19/icpc19.tex
cleaning
intro - not finished

\begin{abstract}
Security is a requirement of utmost importance to produce software with high
quality. However, there is still a considerable amount of zero-day
vulnerabilities being discovered, and fixed, almost weekly. We suspect that when
developers address these vulnerabilities on their codebases, they actually
affect the maintainability of software, potentially introducing other
vulnerabilities. This paper evaluates the impact of refactorings to improve
security on the maintainability of open-source software. Maintainability is
measured based on the \emph{Better Code Hub}'s model of 10 guidelines using a
dataset including $607$ security commits. Results show that fixing software
vulnerabilities hinders the maintainability of open-source software. This is
particularly the case for refactorings that address \emph{Broken Authentication
and Session Management} issues, \emph{Memory Leaks} and \emph{Denial-of-Service}
attacks. Within the implications of our project, we highlight the need for tools 
and documentation to help developers adopt security best practices.
\end{abstract}

@@

Unmaintainable code is hard to test, analyse and re-use. It is hard to find bugs and vulnerabilities in high complex code, this some developers then to forget maintainability. Therefore, we want to bring awareness for the fact that developers need to write maintainable code to prevent the introduction of new vulnerabilities in the code or other defects.

Vulnerabilities in open-source software are similar to exploits found in closed products. These are portions of code that were developed with problems or features that attackers can take advantage to carry out harmful actions. In some cases, these issues are the advantage that hackers need to take services down or even get remote access to others systems. However, whereas closed software is written by developers who are following the same organization rules, open-source is less restrictive and more diverse since it depends on the community who writes, fixes, and generally maintains the codebase.

In 2011, the International Organization for Standardization (ISO) issued an update for software product quality ISO/IEC 25010~\cite{iso:2011} considering \emph{Security} as one of the main software product quality characteristics. In this work, we are concerned about the impact that fixing vulnerabilities may have in the maintainability of open-source software. Are developers affecting software maintainability when trying to improve software safety? We suspect that some of those fixes may have a negative impact in the software maintainability and possibly even be the cause of the introduction of new vulnerabilities.

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to validity of this study.
%
\subsection{Construct}
%
The maintainability formula was inferred based on the reports of the Better Code Hub.
Since the dataset has a broad group of open-source projects with different backgrounds, this metric may not be suitable for all of them. However, BCH does have a representative benchmark of
closed and open-source software projects that is under the calculation of the thresholds used in the computation of each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012} (Table \ref{tab:guidelines}).

@@

All the security commits are described in English.
%
\section{Related Work}\label{sec:rw}

>>>>

\begin{abstract}
Security is a requirement of utmost importance to produce software with high
quality. However, there is still a considerable amount of zero-day
vulnerabilities being discovered, and fixed, almost weekly. We suspect that when
developers address these vulnerabilities on their codebases, they actually
affect the maintainability of software, potentially introducing other
vulnerabilities. This paper evaluates the impact of refactorings to improve
security on the maintainability of open-source software. Maintainability is
measured based on the \emph{Better Code Hub}'s model of 10 guidelines using a
dataset including $607$ security commits. Results show that fixing software
vulnerabilities hinders the maintainability of open-source software. This is
particularly the case for refactorings that address \emph{Broken Authentication
and Session Management} issues, \emph{Memory Leaks} and \emph{Denial-of-Service}
attacks. The observations of this work lead us to conclude that security refactorings should be performed with extra care.
\end{abstract}

@@

Security is a requirement of utmost importance to produce software with high
quality. However, there is still a considerable amount of zero-day
vulnerabilities being discovered, and fixed, almost weekly as disclosed by the Zero Day Initiative\footnote{Zero Day Initiative website available  at https://www.zerodayinitiative.com/advisories/published/ (Accessed on \today{})} website. These vulnerabilities are gateways for attackers to penetrate systems and exploit their resources.

% Maintainability importance is being understimated.

% The time required to implement an additional security functionality in an application is not always easy to estimate. Estimating whether such functionality requires any changes in the architecture of the system is even harder. For instance, if you think about refining the access control policies of your application, by moving from a function-level access control to a data-level access control (where access control is performed on the data field instead on the system functionality executed by the user), can you precisely estimate the amount of time required to perform such change or even if that change has an impact on the architecture of the system?

% Having a resilient and scalable security architecture and design is crucial to prevent vulnerabilities.

% High complex code is hard to test, thus being more difficult to find bugs and vulnerabilities. Therefore, writing maitainable code prevent the introduction of vulnerabilities in the code.

In 2011, the International Organization for Standardization (ISO) issued an update for software product quality ISO/IEC 25010~\cite{iso:2011} considering \emph{Security} as one of the main software product quality characteristics. Unmaintainable code is hard to test, analyse and re-use, and thus more difficult to find vulnerabilities and defects. Therefore, writing maitainable code prevents the introduction of vulnerabilities in the code. In this work, we are concerned about the impact that fixing vulnerabilities may have in the maintainability of open-source software. Are developers affecting software maintainability when trying to improve software safety? We suspect that some of those fixes may have a negative impact in the software maintainability and possibly even be the cause of the introduction of new vulnerabilities.

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to validity of this study.
%
\subsection{Construct}
%
The maintainability formula was inferred based on the reports of the Better Code Hub.
Since the dataset has a broad group of open-source projects with different backgrounds, this metric may not be suitable for all of them. \Sof{Improve it} However, BCH does have a representative benchmark of
closed and open-source software projects that is under the calculation of the thresholds used in the computation of each maintainability guideline (Table \ref{tab:guidelines})~\cite{Visser:2016:OREILLY, Baggen2012}.

@@

All the security commits are described in English. This approach discards security refactorings provided in other languages.

\section{Related Work}\label{sec:rw}
